---
title: "BASIC ANOVA TEST"
author: "DORIS BENIG"
date: "2025-01-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
git init
git remote add origin https://github.com/D4Benig/Stats-R.git
git branch -M main
git add .
git commit -m "Initial commit"
git push -u origin main

```


Does the type of treatment significantly affect the weight of plants?

Hypothesis:

Null Hypothesis $H_0$: The type of treatment does not have a significant effect on the weight of plants.
Alternative Hypothesis $H_1$: The type of treatment has a significant effect on the weight of plants.

# Step 1: Load the necessary libraries 

```{r}
# Data Wrangling libraries 

library(tidyverse)
library(dplyr)
```

# Step 2: Load and explore the datasete

```{r}
# Load the dataset
PlantGrowth
str(PlantGrowth) #This is an R inbuilt dataset
```
The PlantGrowth dataset has 2 variables (weight and group). weight is a numerical variable and group is a factor variable with 3 levels/categories. There are 30 observations in this dataset.  

# Step 3: Visualize the distribution 

The aim is to test if the average weights of plants in the 3 treatment groups are equal or not. 
So, the x-variable is 'group', the categorical variable, and the y-variable is 'weight', the numerical variable. 
Since there is a categorical variable and a numerical variable, visualize the distribution of these variables using a boxplot.

```{r}
ggplot(PlantGrowth, aes(x = group, y = weight))+
  geom_boxplot()+
  labs(title = "Distribution of plant weight ",
       x = "Treatment Groups",
       y = "Plant Weight")
```
Plants treated with Treatment2 has a higher mean weight compared to plants treated with Treatment1 and the Control group. 

# Step 4: Run the ANOVA test

```{r}
summary(aov(PlantGrowth$weight~PlantGrowth$group))

qf(0.95, df1 = 2, df2 = 27)  # 0.95 = 1 - alpha
qf(0.99, df1 = 2, df2 = 27)  # 0.95 = 1 - alpha


```


# Step 5: Interpretation of ANOVA test result 

Suppose $\alpha = 0.01$ , fail to reject the null hypothesis as the p-value from the ANOVA test is 0.0159
The F (2,27) statistic, 4.846 < 5.488118.
There isn't sufficient evidence to conclude that the mean plant weights from the 3 treatment groups are different at 1% significance level. 


Suppose $\alpha = 0.05$ then reject the null hypothesis as the p-value from the ANOVA test is 0.0159 is < 0.05
The F (2,27) statistic, 4.846 > 3.354131 
There is evidence to conclude that the mean plant weights from the 3 treatment groups are different at 5% significance level. 


